The relative roles of genetic, ecological and environmental factors in insular species radiations

### 2018-05-14 paper/draft.tex
remove some space and comments

\begin{abstract}
  Islands are known for their peculiar and fascinating biotas, typically very distinct from any other landmasses, be it neighboring islands or continental areas.
  Although island biogeography theory has seen an ever increasing development over the years,
  the mechanisms underlying these patterns are not fully understood/ not yet explained in a direct causal relationship.
  Specifically, some processes remain un- or under-explored, foremost of these adaption and micro-evolution.
  We use an individual-based mechanistic model implementing genetic and micro-evolutionary processes and other commonly investigated ecological and
  environmental processes to assess the importance of drivers across different organisational and spatio-temporal levels
  on species trait distribution on islands.
  Specifically, we address the question why some lineages diverge stronger than others.
  We found that lineages with intermediate levels of gene linkage evolve into a higher number of distinct populations
  than low or high gene linkage lineages.
  The model enabled us to also take a closer look on the properties of lineages/species that actually managed to colonize islands,
  and which of them resulted in radiations.

\end{abstract}

\section{Introduction}
%% island biota have always fascinated and inspired researchers (Darwin, Wallace, ...)

%% This interest in islands and their recognition as highly suitable model systems resulted in the formulation of
%% important theoretical frameworks, foremost MacArthur and Wilson's seminal equilibrium theory of island biogeography (ETIB).
%% Later ... GDM.

%% These theories have in common a focus on species numbers and respective rates of migration, extinction and speciation.
%% While they are successful in describing species numbers, they are neutral in design (in the UNTB sense) and thus were not meant to consider species traits and compositions.

%% These patterns have been explained theoretically [here, here and here], but usually one at a time.
%% In the following, we aim to produce and thus explain these forementioned patterns using a single framework, and thus following the same set of (rather simple) rules.

%% [NEW ANGLE:]
Islands are understood as suitable model systems for ecology and evolution due to their well definable characteristics like isolation (e.g. Losos \& Ricklefs 2009; Warren et al. 2015  Islands as model systems in ecology and evolution: prospects fifty years after MacArthur‐Wilson. EcoLett). %TODO: ref
Part of this value as study system originates by typically remarkable species. In fact, islands often possess unique species communities and can exceede mainland areas by a factor of 9.5 and 8.1 in plant and animal  endemism, respectively (Kier et al. 2005. A global assessment of endemism and species richness across island and mainland regions. PNAS). %TODO: refs
Such high endemism makes islands  especially threatened by human impacts, in particular climate change and invasive species (Fordham \& Brook 2010). %TODO: refs
Hence, understanding the origin, maintenance, functioning and evolution of island communities is of a high priority objective.

>>>>

\begin{abstract}
  Islands are known for their peculiar and fascinating biotas, typically very distinct from any other landmasses, be it neighboring islands or continental areas.
  Although island biogeography theory has seen an ever increasing development over the years,
  the mechanisms underlying these patterns are not fully understood/ not yet explained in a direct causal relationship.
  Specifically, some processes remain un- or under-explored, foremost of these adaption and micro-evolution.
  We use an individual-based mechanistic model implementing genetic and micro-evolutionary processes and other commonly investigated ecological and
  environmental processes to assess the importance of drivers across different organisational and spatio-temporal levels
  on species trait distribution on islands.
  Specifically, we address the question why some lineages diverge stronger than others.
  We found that lineages with intermediate levels of gene linkage evolve into a higher number of distinct populations
  than low or high gene linkage lineages.
  The model enabled us to also take a closer look on the properties of lineages/species that actually managed to colonize islands,
  and which of them resulted in radiations.
\end{abstract}

\section{Introduction}
Islands are understood as suitable model systems for ecology and evolution due to their well definable characteristics like isolation (e.g. Losos \& Ricklefs 2009; Warren et al. 2015  Islands as model systems in ecology and evolution: prospects fifty years after MacArthur‐Wilson. EcoLett). %TODO: ref
Part of this value as study system originates by typically remarkable species. In fact, islands often possess unique species communities and can exceede mainland areas by a factor of 9.5 and 8.1 in plant and animal  endemism, respectively (Kier et al. 2005. A global assessment of endemism and species richness across island and mainland regions. PNAS). %TODO: refs
Such high endemism makes islands  especially threatened by human impacts, in particular climate change and invasive species (Fordham \& Brook 2010). %TODO: refs
Hence, understanding the origin, maintenance, functioning and evolution of island communities is of a high priority objective.
